---
title: "ACTL4305/5305 Actuarial Data Analytic Application"
author: "Week 1: Data Visualization"
date: "" 
output: 
  bookdown::pdf_book:
    keep_tex: false
    number_sections: yes
    toc: false
bibliography: [refWeek1.bib]
biblio-style: apalike
link-citations: yes
colorlinks: true
fontsize: 10pt
fig_caption: yes
linestretch: 1
geometry: margin=2.3cm
classoption: oneside
graphics: yes
fontfamily: mathpazo
includes:
- \usepackage{float}
- \usepackage{booktabs}
- \usepackage{multicol}
- \floatplacement{figure}{H}
- \usepackage[table]{xcolor}
- \usepackage{amsmath}
- \usepackage{bm}
---

# Learning Objectives {.unnumbered}

-   Learn how to use tidyverse and ggplot() for data visualization.

```{r global-options, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

# Data Visualization

## Data Source

While it might be difficult to obtain data to address a specific research problem or answer a business question, it is relatively easy to obtain data to test a model or an algorithm for data analysis. In the modern era, readers can obtain datasets from the Internet. The following is a list of some websites to obtain real-world data:

-   **UCI Machine Learning Repository.** This website (url: <http://archive.ics.uci.edu/ml/index.php>) maintains more than 400 datasets that can be used to test machine learning algorithms.

-   **Kaggle.** The Kaggle website (url: <https://www.kaggle.com/>) include real-world datasets used for data science competitions. Readers can download data from Kaggle by registering an account.

-   **DrivenData.** DrivenData aims at bringing cutting-edge practices in data science to solve some of the world’s biggest social challenges. In its website (url: <https://www.drivendata.org/>), readers can participate in data science competitions and download datasets.

-   **Analytics Vidhya.** This website (url: <https://datahack.analyticsvidhya.com/contest/all/>) allows you to participate and download datasets from practice problems and hackathon problems.

-   **KDD Cup.** KDD Cup is the annual Data Mining and Knowledge Discovery competition organized by the ACM Special Interest Group on Knowledge Discovery and Data Mining. This website (url: <http://www.kdd.org/kdd-cup>) contains the datasets used in past KDD Cup competitions since 1997.

-   **U.S. Government’s open data.** This website (url: <https://www.data.gov/>) contains about 200,000 datasets covering a wide range of areas including climate, education, energy, and finance.

-   **AWS Public Datasets.** In this website (url: <https://aws.amazon.com/datasets/>), Amazon provides a centralized repository of public datasets, including some huge datasets.

-   **CASdatasets: Insurance Datasets.** In this website (url: <http://cas.uqam.ca/>). A collection of datasets, originally for the book 'Computational Actuarial Science with R' edited by Arthur Charpentier. Now, the package contains a large variety of actuarial datasets.

## Import Data

**CASdatasets** is R friendly, so we can download it by `install.packages("CASdatasets", repos = "http://cas.uqam.ca/pub/", type="source")`. After downloading, when you need this dataset, use `library()` function.

Before you install CASdatasets, make sure you have already installed these packages:

-   install.packages("zoo"),

-   install.packages("xts"),

-   install.packages("sp").

```{r echo=FALSE, warning=FALSE, message=FALSE}
#Before you install CASdatasets, make sure you have already installed these packages:
#install.packages("zoo")

#install.packages("xts")

#install.packages("sp")

#install.packages("CASdatasets", repos = "http://cas.uqam.ca/pub/", type="source")

library(CASdatasets)
#?CASdatasets # Description of this dataset.

data(freMTPL2freq) # import a dataset we want
data(freMTPL2sev) 

attach(freMTPL2freq) # The database is attached to the R search path. This means that the database is searched by R when evaluating a variable, so objects in the database can be accessed by simply giving their names.
attach(freMTPL2sev)
freMTPL2freq$ClaimNb <- as.integer(freMTPL2freq$ClaimNb)
```

Now the datasets called `freMTPL2freq` and `freMTPL2sev` are imported successfully. Have a look at this dataset first. In the two datasets `freMTPL2freq`, `freMTPL2sev`, risk features are collected for 677,991 motor third-part liability policies (observed mostly on one year). In addition, we have claim numbers by policy as well as the corresponding claim amounts. `freMTPL2freq` contains the risk features and the claim number while `freMTPL2sev` contains the claim amount and the corresponding policy ID.

`freMTPL2freq` contains 12 columns:

-   `IDpol`: The policy ID (used to link with the claims dataset).

-   `ClaimNb`: Number of claims during the exposure period.

-   `Exposure`: The period of exposure for a policy, in years.

-   `Area`: The area code.

-   `VehPower`: The power of the car (ordered categorical).

-   `VehAge`: The vehicle age, in years.

-   `DrivAge`: The driver age, in years (in France, people can drive a car at 18).

-   `BonusMalus`: Bonus/malus, between 50 and 350: \<100 means bonus, \>100 means malus in France.

-   `VehBrand`: The car brand (unknown categories).

-   `VehGas`: The car gas, Diesel or regular.

-   `Density`: The density of inhabitants (number of inhabitants per km2) in the city the driver of the car lives in.

-   `Region`: The policy regions in France (based on a standard French classification).

`freMTPL2sev` contains 2 columns:

-   `IDpol` The occurence date (used to link with the contract dataset).

-   `ClaimAmount` The cost of the claim, seen as at a recent date.

## Task 1: How to know the relationship between claim frequency and driver age?

First, we create a figure using the codes from this week's lecture slide.

```{r ageclaimnb, fig.height=4, fig.cap= "DriveAge v.s ClaimNb"}
library(tidyverse)

ggplot(data = freMTPL2freq) + # the dataset
aes(x = DrivAge) + # the x position
aes(y = ClaimNb) +
aes(color = ClaimNb)+
# the y position
geom_point() + # the point geometric shape
# Adjust axis titles' font size
theme(axis.title=element_text(size=14,face="bold"))
```

It seems Figure \@ref(fig:ageclaimnb) is not informative... **WHY?**

In practice, the frequency of most policy is 0. When you create plots in this way, you can see a bunch of data points at bottom level (0). And the number of observations is large (678013), which makes it difficult to recognize the pattern from so many points...

What I will do is plot the average `ClaimNb` for each `DriveAge`. This requires some code you haven't learned in lecture, but it is really helpful in this case.

```{r aveclaimnb, fig.height=4, fig.cap= "DriveAge v.s Average ClaimNB"}
freMTPL2freq %>% 
  group_by(DrivAge) %>% 
  summarize(AveClaimNb=mean(as.double(ClaimNb))) %>% 
  ggplot(aes(x=DrivAge, y=AveClaimNb)) + geom_point() + geom_smooth()
```

From Figure \@ref(fig:aveclaimnb), now we clearly see a pattern! It tells us younger ages less than 25 and older ages around 80 are more likely to make a claim. So, is this the true story?

In general insurance, it's common to calculate $$\text{Claim Rate}_{age}=\cfrac{\sum_i\text{ClaimNb}_{age,i}}{\sum_i \text{Exposure}_{age,i}}$$, where $i$ represents the $i$th policyholder at this $age$. Claim Rate tells us the number of claim per exposure (year), which removes the effect of different exposures. Now let's see the relationship between Claim Rate and `DriveAge`.

```{r aveclaimrate, fig.height=4, fig.cap= "DriveAge v.s Average ClaimRate"}
freMTPL2freq %>% 
  group_by(DrivAge) %>% 
  summarize(claim_rate=sum(as.double(ClaimNb))/sum(Exposure)) %>% 
  ggplot(aes(x=DrivAge, y=claim_rate)) + geom_point() + geom_smooth()
```

From Figure \@ref(fig:aveclaimrate), we can tell that younger ages become more dangerous while older ages are more safe.

Here is another informative figure called Violin plot. What do you learn from this plot?

```{r ViolinDC, fig.height=4, fig.cap= "DriveAge v.s Average ClaimRate"}
# Violin plot DrivAge vs number of claims:
freMTPL2freq %>% filter(ClaimNb<5) %>% 
  mutate(ClaimNb=as.factor(ClaimNb)) %>% 
  ggplot(aes(ClaimNb, DrivAge)) + 
  geom_violin(aes(fill = ClaimNb))
```

## Task 2: How to know the relationship between claim frequency and the area?

Task 2 is a little bit different with Task 1 because the type of `Aera` is factor and the type of `DrivAge` is integer.

```{r}
str(freMTPL2freq)
```

First, we try the code in Task 1. Though we can get some information from Figure \@ref(fig:crarea), it's not straightforward.

```{r crarea, fig.height=4, fig.cap= "Area v.s Average ClaimRate"}
freMTPL2freq %>% 
  group_by(Area) %>% 
  summarise(claim_rate=sum(as.double(ClaimNb))/sum(Exposure)) %>% 
  ggplot(aes(x=Area, y=claim_rate)) + geom_point() + geom_smooth()+
  coord_cartesian(ylim = c(0,0.2)) # set the axis limits
```

A better choice in this case is to use histogram (bar plot) for categorical variable. Now it's clear to compare the Claim Rates in different areas.

```{r crareahis, fig.height=4, fig.cap= "Area v.s Average ClaimRate (histogram)"}
freMTPL2freq %>% # data piped into
group_by(Area) %>% 
summarise(claim_rate=sum(as.double(ClaimNb))/sum(Exposure)) %>% 
ggplot() + # initiating plot
aes(x = Area) + #categorical variable
aes(y = claim_rate) +
geom_col() + #Use `geom_col` to creat a column geometry
aes(color = Area) +
aes(fill = Area) + # new aes 'fill'
aes(linetype = Area)+ #new aes 'linetype'
coord_cartesian(ylim = c(0,0.2)) # set the axis limits
```
